Measuring Subgroup Preferences in Conjoint Experiments

### 2018-12-12 manuscript.Rnw
incorporate changes from Sara 2018-12-12

{\abstract Conjoint analysis is an increasingly prominent tool for studying political preferences. The method powerfully disentangles patterns in respondents' favorability toward complex, multidimensional objects, such as political candidates or public policies. Most conjoint analyses rely upon a fully randomized conjoint design to generate average marginal component effects (AMCEs), which measure the degree to which a given value of a conjoint profile feature increases or decreases respondents' support for the overall profile relative to a baseline, averaging across all respondents and all other profile features. While the AMCE has a clear causal interpretation, most published conjoint analyses also use AMCEs to simply describe preferences (i.e., levels of favorability toward objects), often including comparisons of AMCEs between subgroups of respondents. We show how this descriptive use of these conditional AMCEs can substantially mislead researchers about the degree of agreement or disagreement between subgroups due the simple, familiar, but seemingly forgotten property that interactions are sensitive to the reference category used in regression analysis. We show how researchers can be misled by differences in conditional AMCEs between subgroups, generating inferences about subgroup differences in favorability that have arbitrary sign, size, and significance. We demonstrate the problem using examples drawn from the published literature and provide suggestions for improved reporting and interpretation using two quantities of interest: the marginal mean and the omnibus F-test. Given the rapidly accelerating use of conjoint analyses, these results have critically important implications for research in nearly all areas of political science.}

@@

In these and many other articles, interpretation of these subgroup analyses focus not just on the \textit{causal effects} of profile features within each subgroup (what Hainmueller et al. term ``conditional AMCEs''; 13) but also focus on an implied quantity of interest: the \textit{difference} between two conditional AMCEs across subgroups. Searching for causal effect heterogeneity is an increasingly common feature of experimental analysis \citep{GreenKern2012, RatkovicTingley2017, GrimmerMessingWestwood2017}, yet most conjoint analyses use this difference-in-AMCEs instead to \textit{descriptively} interpret apparent differences in favorability toward objects with a given feature (e.g., immigrants from Syria) between the two groups (e.g., low and high ethnocentrism respondents).

@@

In what follows, we demonstrate the challenges of conjoint analysis and reminder readers of how reference category choice for profile features creates significant problems for comparing conditional AMCEs across respondent subgroups. We show how the use of an arbitrary reference category means the size and the direction of differences-in-AMCEs have little relationship to the underlying degree of favorability of the subgroups toward profiles with particular features and that reference category choices can make similar preferences look dissimilar and dissimilar preferences look similar, using examples drawn from the published political science literature (namely experiments by \citealt{HainmuellerHopkinsYamamoto2014, BallardRosaMartinScheve2016, TeeleKallaRosenbluth2018}). The paper then provides suggestions for improved conjoint reporting and interpretation based around two quantities of interest drawn from the factorial experimentation literature: (1) unadjusted marginal means, a quantity measuring favorability toward a given feature, and (b) an omnibus F-test, measuring differences therein. Newly developed software for the R programming language to support our findings --- and that can be used to examine sensitivity of conjoint analysis to reference category selection, calculate AMCEs and marginal means, perform subgroup analyses, and test for subgroup differences in any conjoint experiment --- is demonstrated throughout \citep{Leeper2018cregg}. We conclude with advice for best practices in the analysis and presentation of conjoint results.

@@

Because levels of features are randomly assigned, pairwise differences between two marginal means for a given feature (e.g., between candidates who are male versus female) have a direct causal interpretation. For fully randomized designs, the AMCE proposed by \citet{HainmuellerHopkinsYamamoto2014} is equivalent to the average marginal effect of each feature level for a model where each feature is converted into a matrix of indicator variables with one level left out as a reference category. This is no different from any other regression context wherein one level of any categorical variable must be omitted from the design matrix in order to avoid perfect multicollinearity.\footnote{In designs that entail constraints between profile features, the average marginal effect is a weighted average of effects across each combination of the constrained features where the weights on the effects are arbitrary but typically uniform. We ignore this distinction in the remainder of this article, as all of our results apply equally to fully randomized and to constrained designs.} This close relationship between AMCEs and marginal means is visible in Figure \ref{fig:hainmueller_candidate_replication} which presents a replication of the AMCE-based analysis of their candidate experiment (upper panel) and an analogous examination of the results using marginal means (lower panel). Here and throughout we use visual presentation of results, but full numerical estimates including appropriate standard errors are presented in the Appendix. Note, in particular, how marginal means convey information about the preferences of respondents for all feature levels while AMCEs definitionally restrict the AMCE for the reference category to zero (or undefined). For example, the AMCE for a candidate serving in the military is 0.09 (or a 9-percentage point) increase in favorability, reflecting marginal means for serving and non-serving candidates of 0.46 and 0.54, respectively.

@@

Being a familiar analytic problem in any regression context, this choice of reference category for estimating AMCEs can seem trivial but is quite consequential. For example, in \citeauthor{HainmuellerHopkinsYamamoto2014}'s candidate experiment, the least liked education level (``no formal education'') is chosen as a reference category, but the authors could have presented the results using any of the categories as the baseline. Figure \ref{fig:reference_category} shows how the estimated AMCEs for each level of education would have differed depending on that choice. Selecting a reference category that receives middling support (i.e., more favorability than some other feature levels but less favorability than others), makes some AMCEs positive and others negative but all AMCEs can be made positive (or negative) simply by choosing a different baseline.\footnote{As another example, in \citeauthor{BallardRosaMartinScheve2016}'s tax preference experiment, the lowest level of taxation is chosen as the reference category for each feature for reasons of substantive interpretation, yet despite the substantive intuitiveness of this, favorability toward the lowest level of taxation is not necessarily higher or lower than preferences for alternative tax rates.} The results would be numerically equivalent --- the alternative linear models used to the estimate the AMCEs have a mathematical equivalence --- but the choice has sizeable consequences for the interpretation of conjoint analyses.\footnote{In \textit{constrained} conjoint designs, the choice of reference category is even more important. Consider, for example, the design of \citeauthor{HainmuellerHopkinsYamamoto2014}'s immigration experiment, which constrains the ``Country of Origin`` feature so that levels `India,' `Germany,' `France,' `Mexico,' `Philippines,' and `Poland' cannot co-occur with the `Escape Persecution' level of the ``Reason for Application'' feature. Consequently, the AMCE for the ``Escape Persecution'' level (relative to the  ``Reunite with family'' reference category) is only defined for the subset of the design involving countries `China,' `Sudan,' `Somalia,' and `Iraq.' The AMCEs for those four countries (relative to India as a baseline) marginalize across all reasons for application, but the AMCEs for the first six countries marginalize only across the latter two reasons. Thus the interpretation of AMCEs --- and the basic ability to estimate them in constrained designs --- depends entirely upon the selection of a reference category where all feature levels can co-occur. In a design where \textit{all} features are constrained, then AMCEs are undefined for the design as a whole and only estimable for subsets of the design that are \textit{conditionally} unconstrained.}

@@

Yet, just as analysis of full sample conjoint data is often descriptive in nature, so too conjoint analysts frequently interpret differences in conditional AMCEs descriptively rather than causally. For example, in one analysis \citet{HainmuellerHopkinsYamamoto2014} visually compare the pattern of AMCEs among high- and low-ethnocentrism respondents and interpret that ``the patterns of support are generally similar for respondents irrespective of their level of ethnocentrism'' (22). \citet{BallardRosaMartinScheve2016} make similar comparisons in their tax policy conjoint: ``While there are few strong differences in preferences for taxing the lower three income groups (the `hard work' group has slightly lower elasticities for taxing the poor), there are strong differences in preferences for taxing the rich. Respondents who believe luck plays a role in economic success are more strongly progressive, although preferences over taxing the \$175K--\$375K bracket are relatively flat'' (12). In these examples, the differences between conditional AMCEs are used as a way of descriptively characterizing differences in \textit{preferences} between the groups rather than differences in \textit{causal effects on preferences} in the groups.

@@

\item Exercise extreme caution when explicitly or implicitly interpreting differences-in-AMCEs across subgroups. While that quantity conveys the difference in effects of a change in a given feature, heterogeneous effects do not necessarily mean different underlying preferences. Differences-in-AMCEs are almost always a biased estimate (of unknown sign and direction) of the difference in underlying preferences. If differences in AMCEs are reported, the choice of reference categories should be discussed explicitly and diagnostics should be provided to justify it.

\item When descriptively characterizing differences in preferences between subgroups, directly estimate the subgroup difference using conditional marginal means and differences between conditional marginal means, rather than relying on difference-in-AMCEs.

@@

\noindent Following this advice, we hope, will allow researchers to more clearly and more accurately represent the results of conjoint experiments.

@@

In an experiment where $N > C$ (the number of observations is large than the number of cells) due to a large sample, or few factors, or levels of each factor, or both (or both of these design characteristic), a sensible next quantity of interest is the \textit{cell mean}: $E[Y|\vec{X} = \vec{x}]$, which in a conjoint simply measures the mean favorability toward a particular profile, $\vec{x}$. An effort to actually estimate this quantity will, however, become obviously intractable when one recognizes that the number of observations in a typical conjoint is much lower the number of feasible profiles ($N \ll C$). The cell mean can be unobserved for many or perhaps most experimental cells.

>>>>

{\abstract Conjoint analysis is an increasingly prominent tool for studying political preferences. The method powerfully disentangles patterns in respondents' favorability toward complex, multidimensional objects, such as political candidates or public policies. Most conjoint analyses rely upon a fully randomized conjoint design to generate average marginal component effects (AMCEs), which measure the degree to which a given value of a conjoint profile feature increases or decreases respondents' support for the overall profile relative to a baseline, averaging across all respondents and all other profile features. While the AMCE has a clear causal interpretation, most published conjoint analyses also use AMCEs to simply describe preferences, often including comparisons of AMCEs between subgroups of respondents. We show how this descriptive use of conditional AMCEs can be substantially misleading about the degree of agreement or disagreement between subgroups due the simple, but often forgotten, property that interactions are sensitive to the reference category used in regression analysis. This leads to inferences about subgroup differences in preferences that have arbitrary sign, size, and significance. We demonstrate the problem using examples drawn from the published literature and provide suggestions for improved reporting and interpretation using two quantities of interest: the marginal mean and the omnibus F-test. Given the rapidly accelerating use of conjoint analyses, this paper makes an important contribution by highlighting pitfalls and presenting advice for best practice in the analysis and presentation of conjoint experiments.}

@@

In these and many other articles, interpretation of these subgroup analyses focus not just on the \textit{causal effects} of profile features within each subgroup (what Hainmueller et al. term ``conditional AMCEs''; 13) but also on an implied quantity of interest: the \textit{difference} between two conditional AMCEs across subgroups. Searching for causal effect heterogeneity is an increasingly common feature of experimental analysis \citep{GreenKern2012, RatkovicTingley2017, GrimmerMessingWestwood2017}, yet most conjoint analyses use this difference-in-AMCEs instead to \textit{descriptively} interpret apparent differences in favorability toward objects with a given feature (e.g., immigrants from Syria) between the two groups (e.g., low and high ethnocentrism respondents).

@@

In what follows, we demonstrate the challenges of conjoint analysis and reminder readers of how reference category choice for profile features creates significant problems for comparing conditional AMCEs across respondent subgroups. We show how the use of an arbitrary reference category means the size and the direction of differences-in-AMCEs have little relationship to the underlying degree of favorability of the subgroups toward profiles with particular features and that reference category choices can make similar preferences look dissimilar and dissimilar preferences look similar, using examples drawn from the published political science literature (namely experiments by \citealt{HainmuellerHopkinsYamamoto2014, BallardRosaMartinScheve2016, TeeleKallaRosenbluth2018}). The paper then provides suggestions for improved conjoint reporting and interpretation based around two quantities of interest drawn from the factorial experimentation literature: (1) unadjusted marginal means, a quantity measuring favorability toward a given feature, and (b) an omnibus F-test, measuring differences therein. Newly developed software for the R programming language to support our findings --- and that can be used to examine sensitivity of conjoint analysis to reference category selection, calculate AMCEs and marginal means, perform subgroup analyses, and test for subgroup differences in any conjoint experiment --- is demonstrated throughout. We conclude with advice for best practices in the analysis and presentation of conjoint results.

@@

Because levels of features are randomly assigned, pairwise differences between two marginal means for a given feature (e.g., between candidates who are male versus female) have a direct causal interpretation. For fully randomized designs, the AMCE proposed by \citet{HainmuellerHopkinsYamamoto2014} is equivalent to the average marginal effect of each feature level for a model where each feature is converted into a matrix of indicator variables with one level left out as a reference category. This is no different from any other regression context wherein one level of any categorical variable must be omitted from the design matrix in order to avoid perfect multicollinearity.\footnote{In designs that entail constraints between profile features, the average marginal effect is a weighted average of effects across each combination of the constrained features where the weights on the effects are arbitrary but typically uniform. We ignore this distinction in the remainder of this article, as all of our results apply equally to fully randomized and to constrained designs.} This close relationship between AMCEs and marginal means is visible in Figure \ref{fig:hainmueller_candidate_replication} which presents a replication of the AMCE-based analysis of the Hainmueller et al. candidate experiment (upper panel) and an analogous examination of the results using marginal means (lower panel). Here and throughout we use visual presentation of results, but full numerical estimates including appropriate standard errors are presented in the Appendix. Note, in particular, how marginal means convey information about the preferences of respondents for all feature levels while AMCEs definitionally restrict the AMCE for the reference category to zero (or undefined). For example, the AMCE for a candidate serving in the military is 0.09 (or a 9-percentage point) increase in favorability, reflecting marginal means for serving and non-serving candidates of 0.46 and 0.54, respectively.

@@

Being a familiar analytic problem in any regression context, this choice of reference category for estimating AMCEs can seem trivial but is quite consequential. For example, in \citeauthor{HainmuellerHopkinsYamamoto2014}'s candidate experiment, the least liked education level (``no formal education'') is chosen as a reference category, but the authors could have presented the results using any of the categories as the baseline. Figure \ref{fig:reference_category} shows how the estimated AMCEs for each level of education would have differed depending on that choice. Selecting a reference category that receives middling support (i.e., more favorability than some other feature levels but less favorability than others), makes some AMCEs positive and others negative but all AMCEs can be made positive (or negative) simply by choosing a different baseline.\footnote{As another example, in \citeauthor{BallardRosaMartinScheve2016}'s tax preference experiment, the lowest level of taxation is chosen as the reference category for each feature for reasons of substantive interpretation, yet despite the substantive intuitiveness of this, favorability toward the lowest level of taxation is not necessarily higher or lower than preferences for alternative tax rates.} The results would be numerically equivalent --- the alternative linear models used to the estimate the AMCEs have a mathematical equivalence --- but the choice has sizeable consequences for the interpretation of conjoint analyses, as we discuss below.\footnote{In \textit{constrained} conjoint designs, the choice of reference category is even more important. Consider, for example, the design of \citeauthor{HainmuellerHopkinsYamamoto2014}'s immigration experiment, which constrains the ``Country of Origin`` feature so that levels `India,' `Germany,' `France,' `Mexico,' `Philippines,' and `Poland' cannot co-occur with the `Escape Persecution' level of the ``Reason for Application'' feature. Consequently, the AMCE for the ``Escape Persecution'' level (relative to the  ``Reunite with family'' reference category) is only defined for the subset of the design involving countries `China,' `Sudan,' `Somalia,' and `Iraq.' The AMCEs for those four countries (relative to India as a baseline) marginalize across all reasons for application, but the AMCEs for the first six countries marginalize only across the latter two reasons. Thus the interpretation of AMCEs --- and the basic ability to estimate them in constrained designs --- depends entirely upon the selection of a reference category where all feature levels can co-occur. In a design where \textit{all} features are constrained, then AMCEs are undefined for the design as a whole and only estimable for subsets of the design that are \textit{conditionally} unconstrained.}

@@

Yet, just as analysis of full sample conjoint data is often descriptive in nature, so too do conjoint analysts frequently interpret differences in conditional AMCEs descriptively rather than causally. For example, in one analysis \citet{HainmuellerHopkinsYamamoto2014} visually compare the pattern of AMCEs among high- and low-ethnocentrism respondents and interpret that ``the patterns of support are generally similar for respondents irrespective of their level of ethnocentrism'' (22). \citet{BallardRosaMartinScheve2016} make similar comparisons in their tax policy conjoint: ``While there are few strong differences in preferences for taxing the lower three income groups (the `hard work' group has slightly lower elasticities for taxing the poor), there are strong differences in preferences for taxing the rich. Respondents who believe luck plays a role in economic success are more strongly progressive, although preferences over taxing the \$175K--\$375K bracket are relatively flat'' (12). In these examples, the differences between conditional AMCEs are used as a way of descriptively characterizing differences in \textit{preferences} between the groups rather than differences in \textit{causal effects on preferences} in the groups.

@@

\item Exercise caution when explicitly or implicitly interpreting differences-in-AMCEs across subgroups. While that quantity conveys the difference in effects of a change in a given feature, heterogeneous effects do not necessarily mean different underlying preferences. Differences-in-AMCEs are almost always a biased estimate (of unknown sign and direction) of the difference in underlying preferences. If differences in AMCEs are reported, the choice of reference categories should be discussed explicitly and diagnostics should be provided to justify it.

\item When descriptively characterizing differences in preferences between subgroups, directly estimate the subgroup difference using conditional marginal means and differences between conditional marginal means, rather than relying on the difference-in-AMCEs.

@@

\noindent Following this advice, we hope, will allow researchers to more clearly and more accurately represent descriptive results of conjoint experiments.

@@

In an experiment where $N > C$ (the number of observations is larger than the number of cells) due to a large sample, or few factors, or levels of each factor, or both (or both of these design characteristic), a sensible next quantity of interest is the \textit{cell mean}: $E[Y|\vec{X} = \vec{x}]$, which in a conjoint simply measures the mean favorability toward a particular profile, $\vec{x}$. An effort to actually estimate this quantity will, however, become obviously intractable when one recognizes that the number of observations in a typical conjoint is much lower the number of feasible profiles ($N \ll C$). The cell mean can be unobserved for many or perhaps most experimental cells.

### 2018-12-12 manuscript.tex
incorporate changes from Sara 2018-12-12

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
## R version 3.5.1 (2018-07-02)
## Platform: x86_64-w64-mingw32/x64 (64-bit)
## Running under: Windows 7 x64 (build 7601) Service Pack 1
## 
## Matrix products: default
## 
## locale:
## [1] LC_COLLATE=English_United Kingdom.1252 
## [2] LC_CTYPE=English_United Kingdom.1252   
## [3] LC_MONETARY=English_United Kingdom.1252
## [4] LC_NUMERIC=C                           
## [5] LC_TIME=English_United Kingdom.1252    
## 
## attached base packages:
## [1] stats     graphics  grDevices utils     datasets  methods   base     
## 
## other attached packages:
## [1] ggplot2_3.1.0 cregg_0.3.0   rio_0.5.11   
## 
## loaded via a namespace (and not attached):
##  [1] zip_1.0.0         Rcpp_1.0.0        cellranger_1.1.0 
##  [4] pillar_1.3.0.9000 compiler_3.5.1    plyr_1.8.4       
##  [7] bindr_0.1.1       forcats_0.3.0     tools_3.5.1      
## [10] digest_0.6.18     lattice_0.20-38   ggstance_0.3.1   
## [13] evaluate_0.12     tibble_1.4.2      gtable_0.2.0     
## [16] pkgconfig_2.0.2   rlang_0.3.0.1     Matrix_1.2-15    
## [19] openxlsx_4.1.0    curl_3.2          haven_1.1.2.9000 
## [22] bindrcpp_0.2.2    withr_2.1.2       stringr_1.3.1    
## [25] dplyr_0.7.8       knitr_1.20        hms_0.4.2        
## [28] lmtest_0.9-36     tidyselect_0.2.5  grid_3.5.1       
## [31] glue_1.3.0        data.table_1.11.8 R6_2.3.0         
## [34] survival_2.43-1   readxl_1.1.0      foreign_0.8-71   
## [37] purrr_0.2.5       magrittr_1.5      splines_3.5.1    
## [40] scales_1.0.0      assertthat_0.2.0  xtable_1.8-3     
## [43] colorspace_1.3-2  sandwich_2.5-0    survey_3.34      
## [46] stringi_1.2.4     lazyeval_0.2.1    munsell_0.5.0    
## [49] crayon_1.3.4      zoo_1.8-4
\end{verbatim}
\end{kframe}
\end{knitrout}

>>>>

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{verbatim}
## R version 3.5.1 (2018-07-02)
## Platform: x86_64-w64-mingw32/x64 (64-bit)
## Running under: Windows 7 x64 (build 7601) Service Pack 1
## 
## Matrix products: default
## 
## locale:
## [1] LC_COLLATE=English_United Kingdom.1252 
## [2] LC_CTYPE=English_United Kingdom.1252   
## [3] LC_MONETARY=English_United Kingdom.1252
## [4] LC_NUMERIC=C                           
## [5] LC_TIME=English_United Kingdom.1252    
## 
## attached base packages:
## [1] stats     graphics  grDevices utils     datasets  methods   base     
## 
## other attached packages:
## [1] ggplot2_3.1.0 cregg_0.3.0   rio_0.5.16   
## 
## loaded via a namespace (and not attached):
##  [1] zip_1.0.0         Rcpp_1.0.0        cellranger_1.1.0 
##  [4] pillar_1.3.0.9000 compiler_3.5.1    plyr_1.8.4       
##  [7] bindr_0.1.1       forcats_0.3.0     tools_3.5.1      
## [10] digest_0.6.18     lattice_0.20-38   ggstance_0.3.1   
## [13] evaluate_0.12     tibble_1.4.2      gtable_0.2.0     
## [16] pkgconfig_2.0.2   rlang_0.3.0.1     Matrix_1.2-15    
## [19] openxlsx_4.1.0    curl_3.2          haven_2.0.0      
## [22] bindrcpp_0.2.2    withr_2.1.2       stringr_1.3.1    
## [25] dplyr_0.7.8       knitr_1.20        hms_0.4.2        
## [28] lmtest_0.9-36     tidyselect_0.2.5  grid_3.5.1       
## [31] glue_1.3.0        data.table_1.11.8 R6_2.3.0         
## [34] survival_2.43-3   readxl_1.1.0      foreign_0.8-71   
## [37] purrr_0.2.5       magrittr_1.5      splines_3.5.1    
## [40] scales_1.0.0      assertthat_0.2.0  xtable_1.8-3     
## [43] colorspace_1.3-2  sandwich_2.5-0    survey_3.34      
## [46] stringi_1.2.4     lazyeval_0.2.1    munsell_0.5.0    
## [49] crayon_1.3.4      zoo_1.8-4
\end{verbatim}
\end{kframe}
\end{knitrout}
